---
title: "Introduction to modelgrid"
author: "Lars Kjeldgaard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to modelgrid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(magrittr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

modelgrid 0.9.0 is now available on CRAN. modelgrid offers a minimalistic but 
very flexible framework to create, manage and train a portfolio of caret models.
Note, you should already be fairly familiar with the caret package before giving
`modelgrid` a spin.

This is the first official release, so below I describe the [key concept](#key_concept)
behind `modelgrid` as well as the features of `modelgrid` divided into three main 
categories:

* [Creating a model grid](#creating_a_model_grid)
* [Training a model grid](#traning_a_model_grid)
* [Editing and removing models from a model grid](#editing_removing_models)

## Key concept behind the model grid {#key_concept}
When facing a Machine Learning problem, you typically want to try out a lot of
models in order to find out, what works and what does not. But how can we manage 
these experiments in a structured, simple and transparent way? You guessed it - 
by using the `modelgrid` package (and yes, I am familiar with the `caretEnsemble`
package, but it was not exactly, what I was looking for). 

A tuning grid consists of combinations of hyperparameters for a specific model. 
A model grid is merely just an extension of that concept in the sense, that it
consists of - potentially many - models, each with their own tuning grid. 
Basically the model grid is built by providing a set of shared settings, that by
default will apply to all models within the model grid, and defining the settings
for the individual models in the model grid.

You can pre-allocate an empty model grid with the constructor function 
`model_grid` and take a look at the structure.

```{r, message = FALSE}
library(modelgrid)
mg <- model_grid()

mg
```

An object belonging to the `model_grid` class has three components:

* `shared_settings`: these are the settings, that will be shared by all models 
in the model grid by default. Generally it will make sense to fix some settings
for all models, e.g. the choice target variable, features, resampling scheme
and sometimes also preprocessing options. By providing them as shared settings
the user avoids redundant code.
* `models`: every individual model specification added to the model grid will be
an element in this list. The individual model specification consists of settings
that uniquely identify the indvidual model. If a setting has been set both as part
of the shared settings and the settings of a given individual model specification,
the setting from the individual model specification will apply for that given
model.
* `model_fits`: this element contains the fitted models (one for each individual
model specification), once the `model_grid` has been trained.

## Creating a model grid {#creating_a_model_grid}

The first natural step of setting up the model grid is to define, what settings that
should be shared by all models by default. We will use the GermanCredit data set
from the caret package as example data and do just that with the `share_settings`
function.

```{r, message = FALSE}
library(magrittr)
library(caret)
library(dplyr)
library(purrr)
# Load data on German credit applications.  
data(GermanCredit)

# Construct empty model grid and define shared settings.
mg <-
  model_grid() %>%
  share_settings(
    y = GermanCredit[["Class"]],
    x = GermanCredit %>% select(-Class),
    preProc = "nzv",
    metric = "ROC",
    trControl = trainControl(
      method = "cv",
      number = 5,
      summaryFunction = twoClassSummary,
      classProbs = TRUE
    )
  )

purrr::map_chr(mg$shared_settings, class)
```

The `shared_settings` component of the model grid is now populated. In order to complete
the model grid we must define a set of individual model specifications, that
we would like to give a shot. A common choice of baseline model could be
a simple parametric model e.g. a Generalized Linear Model. The model specification
is added to the model grid with the `add_model` function.

```{r}
mg <- 
  mg %>%
  add_model(model_name = "Logistic Regression Baseline",
            method = "glm",
            family = binomial(link = "logit"))

mg$models
```

`model_grid` requires a (unique) name for each individual model specification, so I
named this one "Logistic Regression Baseline". If the user does not provide a name,
a generic name - "Model[int]" - is generated automatically. 

This is all it takes to create the smallest possible model grid with only one unique
model configuration. The model grid can be trained with the `train` function, for
more on this go to [Training a model grid](#training_a_model_grid).

But a model grid with only one model specification is obviously not a really interesting
use case. Let us insert two more model specifications into the model grid: another two
logistic regression models, but this time with the features being preprocessed 
with Principal Component Analysis. 

```{r}
mg <- 
  mg %>%
  add_model(model_name = "Logistic Regression PCA",
            method = "glm",
            family = binomial(link = "logit"),
            preProc = c("nzv", "center", "scale", "pca")) %>%
  add_model(model_name = "Logistic Regression PCA 98e-2",
            method = "glm",
            family = binomial(link = "logit"),
            preProc = c("nzv", "center", "scale", "pca"),
            custom_control = list(preProcOptions = list(thresh = 0.98)))
            
mg$models
```

You can of course add as many models as you like to the model grid with
the `add_model` function. 

## Training a model grid {#training_a_model_grid}

The models from a model grid can be trained with the `train` function 
from the `caret` package, for which I have implemented a S3 method for the
`model_grid` class.

When you call `train` with a `model_grid`, all of the individual model
specifications are [consolidated](#consolidation) with the shared settings into complete
caret model specifications, which are then [trained](#train) one by one with
caret.

### Consolidation of settings into complete caret model specifications {#consolidation}

For a given model the model settings are consolidated with the
`consolidate_model` function. Let us see how this works with the three models.
For the baseline model there is no overlap between the shared settings and
the settings in the individual model spec, and the settings will just 
be appended into one configuration.

```{r}
# there are no conflicts.
dplyr::intersect(names(mg$shared_settings), names(mg$models$`Logistic Regression Baseline`))

# consolidate model settings into one model.
consolidate_model(
  mg$shared_settings, 
  mg$models$`Logistic Regression Baseline`
  ) %>%
  purrr::map_chr(class)
```

In case the same setting has been specified both in the shared settings of the
model grid and in the individual settings for a specific model, the individual
setting will apply. This is the case for the model "Logistic Regression PCA", 
where the 'preProc' argument has also been defined in the model specific
configuration.

```{r}
# the 'preProc' setting is defined both in the shared and model specific settings.
dplyr::intersect(names(mg$shared_settings), names(mg$models$`Logistic Regression PCA`))

mg$shared_settings$preProc
mg$models$`Logistic Regression PCA`$preProc

# consolidate model settings into one model.
consolidate_model(
  mg$shared_settings, 
  mg$models$`Logistic Regression PCA`
  ) %>%
  magrittr::extract2("preProc")
```

Also, if the 'trControl' argument is defined as part of the shared settings, the
subsettings of 'trControl' can be modified for a specific model with the special
setting 'custom_control' (which itself is given as an explicit argument to the 
`add_model` function) in the model specific settings.

For the model `Logistic Regression PCA 98e-1`, the preprocessing options for PCA
were adjusted with 'custom_control'. When the model is consolidated, the model 
specific customizations of subsettings of the shared 'trControl' argument will
apply.

```{r}
# the 'trControl$preProcOptions$thresh' setting is defined in the shared
# settings but customized in the model specific settings.

mg$shared_settings$trControl$preProcOptions$thresh
mg$models$`Logistic Regression PCA 98e-2`$custom_control$preProcOptions$thresh

# consolidate model settings into one model.
consolidate_model(
  mg$shared_settings, 
  mg$models$`Logistic Regression PCA 98e-2`
  ) %>%
  magrittr::extract2(c("trControl", "preProcOptions", "thresh"))
```

### Model training {#train}

When calling the `train` function, the `consolidate_model` function is called
under the hood with all of the individual models and the shared settings, and
a set of complete caret model specifications is generated - one for each 
individual model specification.

Afterwards the models are trained one by one with `caret`, and the fitted 
models are saved in the `model_fits` component of the model grid.

```{r, message = FALSE, warning = FALSE}
# train models from model grid.
mg <- train(mg)

# the fitted models now appear in the 'model_fits' component.
names(mg$model_fits)

# extract performance.
mg$model_fits %>%
  caret::resamples(.) %>%
  summary(.)
```

If we now add an additional models to the model grid, and call train on the model
grid again, only the new models (those that do not yet have a fit) will be trained
by default. 

```{r, warning = FALSE}
# train models from model grid.
mg <- 
  mg %>%
  add_model(model_name = "Funky Forest",
            method = "rf") %>%
  train(.)

names(mg$model_fits)
```

If you call `train` with the `train_all` argument set to TRUE, all models will
be trained regardless.

### Support for all train interfaces

The training of a `model_grid` supports both the explicit 'x', 'y' interface to train, 
the formula interface and last but not least the new powerful 'recipe' interface.
Let us try out the latter. First we will create a base recipe.

```{r, message = FALSE}
# create base recipe.
library(recipes)
rec <- 
  recipe(GermanCredit, formula = Class ~ .) %>%
  step_nzv(all_predictors())
```

With that as a starting point I will create and train a minimal model grid 
as an example. I will tweak the recipe for one of the models.

```{r, warning = FALSE}
mg_rec <-
  model_grid() %>%
  share_settings(
    metric = "ROC",
    data = GermanCredit,
    trControl = trainControl(
      method = "cv",
      number = 5,
      summaryFunction = twoClassSummary,
      classProbs = TRUE
    )
  ) %>%
  add_model(
    model_name = "Log Reg",
    x = rec,
    method = "glm",
    family = binomial(link = "logit")
  ) %>%
  add_model(
    model_name = "Log Reg PCA",
    x = rec %>%
      step_center(all_predictors()) %>%
      step_scale(all_predictors()) %>%
      step_pca(all_predictors()),
    method = "glm",
    family = binomial(link = "logit")
  ) %>%
  train(.)

mg_rec$model_fits %>%
  caret::resamples(.) %>%
  summary(.)
```

## Editing and removing models {#editing_removing_models}

`modelgrid` has a couple of functions, that makes it easy to work iteratively
with the model specifications in a model grid. If you want to modify an
existing model configuration, please use the `edit_model` function. Below
I use it to modify one of the GLM models.

```{r}
# existing model configuration.
mg$models$`Logistic Regression PCA`

# edit model configuration.
mg <-
  mg %>%
  edit_model(model_name = "Logistic Regression PCA",
             preProc = c("nzv", "center", "scale", "ICA"))

mg$models$`Logistic Regression PCA`
```

As you see, when you modify an existing model specification, any corresponding
fitted model is deleted

You can also remove a model specification (including any fitted model) from 
the model grid with the `remove_model` function.

```{r}
names(mg$models)

# remove model configuration.
mg <-
  mg %>%
  remove_model("Funky Forest")

names(mg$models)
```








